Mining Scenario/Transaction Pairs HW/SW Co-verification Using Symbolic Execution

### 2017-04-03 paper.tex
Approach 1 description complete

%
%===============================================================================
\section{Introduction}
%===============================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Pre-RTL versus Post-RTL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Hardware/software co-verification is broadly performed at two different
phases in the design cycle: pre-RTL and post-RTL.  In the pre-RTL phase,
engineers write high-level software (SW) models for Intellectual Property (IP), 
which are commonly known as {\em Transaction Level Model}s
(TLM)~\cite{codes14}.  TLMs are typically implemented using the SystemC TLM
library~\cite{tlm} or as plain C programs.  The goal of a TLM is to capture
enough functionality of the hardware (HW) such that it is possible to execute the
SW (drivers/firmware) against it.  Consequently, the TLMs are
incomplete, as they are designed to contain only those behaviors that are
observable by the SW (the drivers or firmware).  In particular, TLMs
are often not cycle-accurate.  These models are used to verify high-level
functionality of the system well before the RTL is
available~\cite{codes14,codes15}. Co-verification of the TLM and the
SW is typically performed by means of testing.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Challenges in Co-verification 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
However, the verification of FW given in C/C++ together with the RTL given in a hardware
description language (HDL) such as Verilog/VHDL is a challenging task due to 
few reasons -- 1) the timing and semantic mismatch between the synchronous
clock-driven HW model and asynchronous event-driven SW model -- for example,
FW running on a processor could be much faster than the HW
model it interacts with, 2) lack of co-specification model that precisely
captures the design intent of the interaction behavior, 3) lack of automated
formal tools that support co-designs that are mostly implemented in C and 
Verilog/VHDL.   In this paper, we address some of these issues to facilitate 
wider adoption of formal techniques in hardware/software co-verification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                     Handling Semantic Gap  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Generating a unified FW-HW model~\cite{fmsd02,emsoft07,fase10} for 
co-verification is a well-known phenomenon to handle the semantic mismatch 
between the HW~\footnote{Note that whenever we refer to ``hardware'' from this point onwards, we mean an RTL implementation 
and not a TLM.}  and FW models.  Most of the existing techniques are not 
useful in practise for three reasons -- 1) use of abstract HW models 
instead of register-transfer level implementation, 2) the HW-FW model
construction is agnostic to the underlying reasoning engine, 3) does not
leverage the co-specification model, hence ineffective for practical
co-verification.  We will call the unified FW-HW model as \emph{co-verification model}.

In this paper, we present two different approaches for automatic formal
verification of HW/SW co-designs -- \emph{monolithic} and \emph{path-based}.  
We embody these techniques in our tools, \hwcbmcv and \verifox, respectively. 
A desirable characteristic of a co-verification model is the ability to exploit 
the expressivity of the underlying reasoning engines, for example, constructing 
a bit-level co-verification model for SAT solvers or a word-level model for SMT 
solvers.  \hwcbmcv constructs a co-verification model through 
in tandem symbolic execution of the HW and SW models.  To this end, a bit-level 
netlist (typically represented as And Inverter Graph) or word-level netlist 
(represented in tool specific format) is synthesized from a HW Register Transfer 
Level (RTL) implementation. Similarly, the SW, given in C, is automatically 
translated into a Static Single Assignment (SSA) form.  A monolithic verification 
condition is generated from these two models which is then solved using a 
satisfiability solver.  Whereas, a co-verification model in \verifox is constructed 
by first translating the RTL description into a software representation in C followed 
by sequentializing the hardware and firmware models.  This results in a single
threaded sequential co-verification model which is represented in C. A path-wise 
symbolic execution is then performed over the co-verification model to
effectively mine only relevant interaction behaviors between the SW and HW
models.

%A key to building an effective co-verification tool is to translate the HW
%and SW models to a common formal semantics called, 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                     Co-specification Model  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A co-verification models must leverage the specification of system-level 
properties across HW/SW boundaries for uncovering critical co-design bugs 
and for effective co-verification.  Our co-specification models are 
specified in C with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. Our
co-verification models in \hwcbmcv and \verifox can be easily verified against 
these co-specification models. The co-design properties expressed by 
co-specification models are sufficiently expressive, amenable to 
verification reuse and can be readily used in assertion-based verification.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %        Difference between Monolithic and Path-wise Approach
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A monolithic approach in \hwcbmcv based on Bounded Model Checking~\cite{CKY03}, 
cannot exploit this infeasibility.  Whereas, a path-based symbolic execution in 
\verifox can automatically infer such Scenario/Transaction pairs and check their 
validity using backend solvers.  For building effective and scalable formal 
co-verification tool, it is therefore imperative to automatically mine only 
relevant scenario/transaction pairs that occurs in practical co-designs.

@@

\subsection{Monolithic Co-verification Approach}

\subsection{Path-based Co-verification Approach}
%
Figure~\ref{proposed-flow} describes our overall tool framework. In the
subsequent section, we describe each step in details.

\begin{figure}[t]
%\centering
\scalebox{.48}{\import{figures/}{work-flow.pspdftex}}
\caption{Proposed HW/SW Co-verification Flow
\label{proposed-flow}}
\end{figure}

%\subsection{Software netlist}
{\bf Software netlist:}
The hardware circuits in Verilog RTL is automatically 
synthesized to a word level C program, called a 
\textit{software netlist}~\cite{mtk2016,mskm2016} using the 
tool {\em v2c}~\cite{mtk2016}. {\em v2c} follows synthesis semantics and not 
simulation semantics. The generated code is cycle-accurate as well as 
bit-precise. {\em v2c} automatically translates concurrent statements 
in Verilog such as procedural blocks and continuous assignments to 
a semantically equivalent C program.

A \textit{software netlist} $\mathcal{SN}$ is defined as follows:
\Omit{
a four-tuple $\langle L, A, l_0, l_e
\rangle$, where $L$ is a finite set of locations, modelling the values of
the program counter in the corresponding sequential code, $l_0 \in L$ is the initial
location, $l_e \in L$ is the error location, and $A \subseteq L \times M
\times L$ is the control flow automaton.  The edges in $A$ are labelled
with a quantifier-free first-order formula over program variables drawn from
the set $M$. These formulas encode an assignment and
take the form $\langle \mathit{In}, \mathit{Out},
\mathit{Seq}, \mathit{Comb}, \mathit{Asgn} \rangle$,
\sjcmt{$M$ is confusing. Is it a set of first order formula
or a set of $Asgn$?}
}
%
\[ 
\begin{array}[t]{@{}lll}
\mathcal{SN} & {:}{:}{=} & \mathit{Modules} \\
\mathit{Modules} & {:}{:}{=} & \epsilon \mid \mathit{moduleName(Var_1,\dots,Var_n)} \{ Asgn \} \;\; Modules \\
\mathit{Asgn} &  {:}{:}{=} & \mathit{CAsgn} \mid \mathit{SAsgn} | moduleName(Var_1,\dots,Var_n)\\
\mathit{CAsgn} & {:}{:}{=} & (V_c := Bv) \mid (V_c := \mathit{Bool}), \quad V_c \in \mathit{Comb} \uplus \mathit{Out} \\
\mathit{SAsgn} & {:}{:}{=} & (V_s := Bv) \mid (V_s := \mathit{Bool}), \quad V_s \in \mathit{Seq} \\
\mathit{Bv} &  {:}{:}{=} & bv_{const} \mid bv_{var} \mid
	\mathit{ITE}(Bool, Bv, Bv) \mid
bv_{op}(bv_1, \ldots, bv_n), \\
& & bv_i \in Bv, \\ 
& & bv_{op} \in \{n-\text{ary operators over bitvectors} \} \\
\mathit{Bool} & {:}{:}{=} & \mathit{true} \mid \mathit{false} \mid \neg{Bool} \mid Bool \land Bool \mid 
Bool \lor Bool \mid \\ 
& & Bv\; relop\; Bv, relop \in \{<,\leq,=,>,\geq\} \\
\end{array}
\]
%
Here, $\mathit{In}$, $\mathit{Out}$, $\mathit{Seq}$, $\mathit{Comb}$
are programs variables that model \emph{input}, \emph{output}, 
\emph{sequential/state-holding} and \emph{combinational/stateless} 
signals in the circuit.  $Asgn$ is a finite set of assignments 
to $Out$, $Seq$ and $Comb$.

\Omit{
A state transition in HW can be described by a set of register updates,
defined by the next-state function, and assignment of non-deterministic 
values to external inputs. The state transition in the software netlist 
corresponds to updates to the variables in $Seq$ and explicit assignment 
of non-deterministic values to the input variables, $In$.
}

%\subsection{Generating Unified FW-HW model}
\subsection{Analyzing FW/HW interaction Patterns}
%{\bf Analyzing FW/HW interaction pattern: }
The data dependencies between FW and HW is analyzed 
to reduce the number of interleavings between them and 
generate a unified FW-HW model. 
\Omit{
Recall that a HW transaction $(\mathcal{HT})$ is in response to a FW transaction 
$(\mathcal{FT})$. Similarly, $\mathcal{FT}$ is in response to the 
inputs from $\mathcal{HT}$ or operating system or application level software. 
The specific interaction pattern between $\mathcal{FT}$ 
and $\mathcal{HT}$ is exploited to generate a 
unified FW-HW model, $\mathcal{PN}$, in C. 
}
We observe that this interaction pattern follows 
producer-consumer relationship as shown in 
Figure~\ref{fig:interleaving1}. 
\Omit{
In practice, a FW thread 
is mostly independent of the HW thread they interact with~\cite{hvc}.  
Specifically, a FW thread is only responsible for configuring the 
HW transactions by writing to memory mapped registers 
or polling the interrupt status register for data transmission or 
receiving incoming data packets. 
}
We formally define the data dependencies between $\mathcal{FT}$ 
and $\mathcal{HT}$. A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.

\Omit{
\textbf{Producer:} 
$\mathcal{FT}$ (or $\mathcal{HT}$) is called a {\em producer} 
if it produces a data ($d$) that is independent of $\mathcal{HT}$ (or $\mathcal{FT}$). 
Thus, the interleavings between $\mathcal{FT}$ and $\mathcal{HT}$ 
can be avoided to compute $d$. 
}

\textbf{Producer-consumer Interaction:}
%
An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.

\Omit{
shared register only FT is writing 
shared register that HT is writing 
and these two sets are disjoint
$\mathcal{FT}$ and $\mathcal{HT}$ is said to exhibit   
{\em producer-consumer} interaction pattern specific 
to data $d$ if $\mathcal{FT}$ produces $d$ and $\mathcal{HT}$ 
consumes $d$. Similarly, when $\mathcal{HT}$
is a producer of $d$ and $\mathcal{FT}$ is a consumer of $d$, 
$\mathcal{HT}$ and $\mathcal{FT}$ is said to exhibit   
{\em producer-consumer} interaction pattern specific to $d$. 
A producer-consumer interaction is defined with respect 
to the data produced. 
}

Figure~\ref{fig:interleaving1} illustrates the FW-HW interaction pattern 
for the UART IP. The FW CFG is shown in the left and the 
HW RTL code snippet is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for our illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$. These interaction patterns 
between $\mathcal{HT}$ and $\mathcal{FT}$ is exploited to generate a 
sequential unified {\em program netlist} model which is defined next.\\ \\

@@

\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}

@@

{\bf Unified Co-verification model:}
A {\em program netlist} $\mathcal{PN}$ is a sequential unified
model of FW and HW. $\mathcal{PN}$ is formally defined as follows.

@@

\mathcal{PN} & {:}{:}{=} & \mathit{Procedure} \\

@@

over variables $Var$ of $\mathcal{PN}$ such that $(exp \in Exp)$ and 

@@

\Omit{
$\mathcal{PN}$ is verification equivalent 
to the concurrent execution of FW and $\mathcal{SN}$.
}
\textbf{Property Driven Slicing of Program Netlist:}
A property-driven slicing is performed on the unified model, $\mathcal{PN}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration. 
\Omit{
\section{Extracting relevant FW/HW interactions using}
Our unified FW-HW model, $\mathcal{PN}$, is composed of a set of 
FW scenarios, $\{\mathcal{S}\}$, and a set of HW transactions, $\{\mathcal{T}\}$ 
that $\{\mathcal{S}\}$ depends on. However, only a subset of $\{\mathcal{T}\}$ is 
relevant for a particular $\mathcal{S}_i \in \mathcal{S}$. In this section, 
we present a path-based exploration algorithm to automatically infer 
interacting $\mathcal{S}/\mathcal{T}$ pairs from $\mathcal{PN}$. 
}
%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
%\Omit{
%\begin{figure}[tp]
%\vspace*{0.3cm}
%\scalebox{.50}{\import{figures/}{work-flow.pspdftex}}
%\caption{Automated Slicing and Trace partitioning\label{fig:desc}}
%\end{figure}
%}
\Omit{
\tikzstyle{rblock}=[rounded corners,draw=black]
\tikzstyle{block}=[draw=black]
\begin{tikzpicture}

@@

\end{tikzpicture}
\caption{Automated Slicing and Trace partitioning\label{fig:desc}}
\end{figure}
}

\input{verifox}
\Omit{
\rmcmt{
The hardware and firmware run independently of each other. 
The communication between them takes place through interrupts. 
The interrupts are modelled by function calls. 
}
}
%
\section{The Conventional Approach to Co-verification}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{Conventional HW/SW Co-verification Flow}
\label{fig:conventional}
\end{center}
\end{figure}

Figure~\ref{fig:conventional} illustrates the conventional approach to HW/SW
co-veri\-fi\-cation, as implemented in the state-of-the-art co-verification
tool, \hwcbmcv~\cite{CKY03,vlsid13}.  \hwcbmcv is the only open-source
formal co-verification tool that supports Verilog RTL (System Verilog 2009
standard) and C/C++ firmware.  In contrast to our proposed approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  The bottom flow illustrates the
translation of the C program into static single assignment (SSA) form. 
These two flows meet only at the solver.  Thus, \hwcbmcv generates a
monolithic formula from the C and RTL description, which is then checked
with SAT/SMT solvers.  \hwcbmcv provides specific handshake primitives such
as $next\_timeframe()$ and $set\_inputs()$ to model FW-HW communications.
%
\Omit{
The hardware and firmware run independently of each other. 
The communication between them takes place through these function calls. 
}

%===============================================================================
\section{Experimental Results}
%===============================================================================

>>>>

%
%===============================================================================
\section{Introduction}
%===============================================================================
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Pre-RTL versus Post-RTL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Hardware/software co-verification is broadly performed at two different
phases in the design cycle: pre-Register Transfer Level (pre-RTL) and 
post-RTL.  In the pre-RTL phase,
engineers write high-level software (SW) models for Intellectual Property (IP), 
which are commonly known as {\em Transaction Level Model}s
(TLM)~\cite{codes14}.  TLMs are typically implemented using the SystemC TLM
library~\cite{tlm} or as plain C programs.  The goal of a TLM is to capture
enough functionality of the hardware (HW) such that it is possible to execute the
SW (drivers/firmware) against it.  Consequently, the TLMs are
incomplete, as they are designed to contain only those behaviors that are
observable by the SW (the drivers or firmware).  In particular, TLMs
are often not cycle-accurate.  These models are used to verify high-level
functionality of the system well before the RTL is
available~\cite{codes14,codes15}. Co-verification of the TLM and the
SW is typically performed by means of testing.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                    Challenges in Co-verification 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The verification of FW given in C/C++ together with the RTL given in a hardware
description language (HDL) such as Verilog/VHDL is a challenging task due to 
few reasons -- 1) the timing and semantic mismatch between the synchronous
clock-driven HW model and asynchronous event-driven SW model -- for example,
FW running on a processor could be much faster than the HW
model it interacts with, 2) lack of co-specification model that precisely
captures the design intent of the communication between HW and FW, 3) lack 
of automated formal tools that support co-designs implemented in C/C++ and 
Verilog/VHDL.   In this paper, we address some of these challenges to enable
seamless adoption of formal techniques in hardware/software co-verification.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                     Handling Semantic Gap  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Generating a unified FW-HW model for co-verification is a 
well-known phenomenon to handle the semantic mismatch between the 
HW~\footnote{Note that whenever we refer to ``hardware'' 
from this point onwards, we mean an RTL implementation 
and not a TLM.}  and FW models.  Most of the existing techniques 
in~\cite{fmsd02,emsoft07,fase10} for constructing a unified FW-HW model are not 
useful in practise for three reasons -- 1) use of abstract HW models 
instead of register-transfer level implementation, 2) the HW-FW model
construction is agnostic to the underlying reasoning engine, 3) does not
leverage the co-specification model, hence ineffective for practical
co-verification.  We will call the unified FW-HW model as \emph{co-verification model}.

In this paper, we present two different approaches for automatic formal
verification of HW/SW co-designs -- \emph{monolithic} and \emph{path-based}.  
We embody these techniques in our tools, \hwcbmcv and \verifox, respectively. 
A desirable characteristic of a co-verification model is the ability to exploit 
the expressivity of the underlying reasoning engines, for example, constructing 
a bit-level co-verification model for SAT solvers or a word-level model for SMT 
solvers.  To this end, \hwcbmcv constructs a co-verification model through 
in tandem symbolic execution of the HW and FW models.  A bit-level 
netlist representation in And Inverter Graph (AIG) format or word-level netlist in
format similar to SMT-LIB standards is synthesized from a HW Register Transfer 
Level (RTL) implementation. Similarly, the SW, given in C, is automatically 
translated into a Static Single Assignment (SSA) form.  A monolithic verification 
condition is generated from these two models which is then solved using a 
satisfiability solver.  Whereas, a co-verification model in \verifox is constructed 
by first translating the RTL description into a software representation in C followed 
by sequentializing the hardware and firmware models.  This results in a single
threaded sequential co-verification model which is represented in C. A path-wise 
symbolic execution is then performed over the co-verification model to
automatically mine relevant interaction behaviors between the FW and HW models.

%A key to building an effective co-verification tool is to translate the HW
%and SW models to a common formal semantics called, 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %                     Co-specification Model  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A co-verification models must enable the specification of system-level 
properties across HW/SW boundaries for uncovering critical co-design bugs 
and for effective co-verification.  Our co-specification models are 
specified in C with the support of special primitive, called {\em next\_timeframe()}, 
that can express temporal correlation between HW and FW events. The
co-verification models in \hwcbmcv and \verifox can be easily verified against 
the co-specification models.  The co-design properties expressed by 
co-specification models are sufficiently expressive, amenable to 
verification reuse and can be readily used in assertion-based verification.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %        Difference between Monolithic and Path-wise Approach
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A monolithic approach in \hwcbmcv is based on Bounded Model Checking~\cite{CKY03}, 
hence it cannot exploit the infeasibility arising from scenario driven constrained 
interaction behaviors.  Whereas, a path-based symbolic execution in 
\verifox can automatically infer such Scenario/Transaction pairs and check their 
validity using backend solvers.  We experimentally show that for building 
effective and scalable formal co-verification tool, it is imperative to 
automatically mine only relevant scenario/transaction pairs that occurs in 
practical co-design environment.

@@

\subsection{Approach 1: Path-based Co-verification Technique}
Figure~\ref{proposed-flow} describes our overall tool framework. In the
subsequent section, we describe each step in details.\\ 

@@

\scalebox{.48}{\import{figures/}{work-flow.pspdftex}}
\caption{Proposed HW/SW Co-verification Flow
\label{proposed-flow}}

@@

\textbf{Step 1: RTL to Software}
A hardware circuit in Verilog RTL is automatically 
translated into a word level C program, called a 
\textit{software netlist}~\cite{mtk2016,mskm2016,mkm2015} following synthesis 
semantics. The software netlist model is cycle-accurate as well as 
bit-precise~\footnote{http://www.cprover.org/hardware/v2c/}. The 
translation automatically handles concurrent statements 
in Verilog such as procedural blocks and continuous assignments, multiple always
blocks, multiple modules and translates them into an equivalent C program.  This
is a radically new but promising approach for synthesizing software 
models from RTL description, mainly targeted for word-level formal hardware 
verification using software verification techniques such as abstract 
satisfaction~\cite{leo-thesis}.  To the best of our knowledge, this is the first 
attempt that demonstrate the efficacy of such translation in the context of 
hardware/software co-verification. 
A \textit{software netlist} $\mathcal{SN}$ is defined as follows:

@@

\mathcal{SN} & {:}{:}{=} & \mathit{Modules} \\
\mathit{Modules} & {:}{:}{=} & \epsilon \mid \mathit{moduleName(Var_1,\dots,Var_n)} \{ Asgn \} \;\; Modules \\
\mathit{Asgn} &  {:}{:}{=} & \mathit{CAsgn} \mid \mathit{SAsgn} | moduleName(Var_1,\dots,Var_n)\\
\mathit{CAsgn} & {:}{:}{=} & (V_c := Bv) \mid (V_c := \mathit{Bool}), \quad V_c \in \mathit{Comb} \uplus \mathit{Out} \\
\mathit{SAsgn} & {:}{:}{=} & (V_s := Bv) \mid (V_s := \mathit{Bool}), \quad V_s \in \mathit{Seq} \\
\mathit{Bv} &  {:}{:}{=} & bv_{const} \mid bv_{var} \mid
	\mathit{ITE}(Bool, Bv, Bv) \mid
bv_{op}(bv_1, \ldots, bv_n), \\
& & bv_i \in Bv, \\ 
& & bv_{op} \in \{n-\text{ary operators over bitvectors} \} \\
\mathit{Bool} & {:}{:}{=} & \mathit{true} \mid \mathit{false} \mid \neg{Bool} \mid Bool \land Bool \mid 
Bool \lor Bool \mid \\ 
& & Bv\; relop\; Bv, relop \in \{<,\leq,=,>,\geq\} \\
Here, $\mathit{In}$, $\mathit{Out}$, $\mathit{Seq}$, $\mathit{Comb}$
are programs variables that model \emph{input}, \emph{output}, 
\emph{sequential/state-holding} and \emph{combinational/stateless} 
signals in the circuit.  $Asgn$ is a finite set of assignments 
to $Out$, $Seq$ and $Comb$. 
\textbf{Step 2: Constructing Co-verification Model}
An unified co-verification model is constructed by sequentializing 
HW and its interacting FW threads, thereby generating a sequential 
single-threaded co-verification model in C.  The motivation behind
sequentialization is based on our observation that practical industrial
co-designs exhibit producer-consumer interaction behavior.  We observe 
that in practise, a FW thread is mostly independent of the HW thread they interact 
with~\cite{hvc,codes14}.  This behavior is also prevelant in industrial co-designs, 
such as IBM Text Accelerator co-design~\cite{polig2014micro,polig2014fpl}
and IBM coherent FFT coprocessor~\cite{giefers2015accelerating}.
Specifically, a FW thread is only responsible for configuring the 
HW transactions by writing to memory mapped registers or polling 
the interrupt status register for data transmission or receiving 
incoming data packets. This interaction pattern exhibits producer-consumer 
relationship, which is also evident in case of our UART example in 
Figure~\ref{fig:interleaving1}.  
We formally define a single-threaded sequential unified co-verification model, 
denoted by $\mathcal{M}_{seq}$. 
  \mathcal{M}_{seq} & {:}{:}{=} & \mathit{Procedure} \\

@@

over variables $Var$ of $\mathcal{M}_{seq}$ such that $(exp \in Exp)$ and 

@@

%\textbf{Producer-consumer Interaction:}
A communicating pair of executions of 
$\mathcal{FT}$ and $\mathcal{HT}$ is an {\em interaction}, which 
takes place through shared variables (or shared registers)
only.  An interaction is said to exhibit {\em producer-consumer}
interaction pattern if 1) the set of variables $V$ written by 
the execution of $\mathcal{FT}$ and and the set of variables 
$V'$ written by the execution of $\mathcal{HT}$ are disjoint 
($V \cap V' = \emptyset$), (2) and the executions 
of $\mathcal{FT}$ and $\mathcal{HT}$ consumes values 
in $V'$ and $V$, respectively.

Figure~\ref{fig:interleaving1} illustrates the FW-HW interaction pattern 
for the UART IP. The FW CFG is shown in the left and the 
HW RTL code snippet is shown in the right. We abuse notations and 
represent $\mathcal{HT}$ in the form of RTL and $\mathcal{FT}$ as CFG
for our illustration purpose. Here, $\mathcal{HT}$ is the producer of interrupt 
status signal, $irq\_o$, which is read by $\mathcal{FT}$ in state $s_1$. 
The signal $irq\_o$ is  triggered when transmitter buffer is 
{\em EMPTY} or receiver buffer is {\em FULL}. 
$\mathcal{HT}$ also produces interrupt status by writing to the 
memory mapped register, $UART\_IS$, which is consumed by 
$\mathcal{FT}$ in state $s_2$.  In another interaction, $\mathcal{HT}$
sends a data-packet to $UART\_TR$ which is consumed by $\mathcal{FT}$ in state $s_7$.
Similarly, $\mathcal{FT}$ is the producer of $data$, $(outb(data,UART\_TR))$, 
in state $s_6$ which is consumed by $\mathcal{HT}$. These interaction patterns 
between $\mathcal{HT}$ and $\mathcal{FT}$ is exploited to generate a 
sequential unified {\em program netlist} model which is defined next.\\ \\
%
\begin{figure}[t]
\scalebox{.55}{\import{figures/}{interleaving.pspdftex}}
\caption{Producer Consumer Interactions Pattern
\label{fig:interleaving1}}
\end{figure}
%
\textbf{Step 3: Property Driven Slicing of Co-verification Model}
A property-driven slicing is performed on the unified model, $\mathcal{M}_{seq}$.  
This step is purely syntactic, meaning that we perform a backward dependency 
analysis starting from the property which only preserve those 
program fragments that are relevant to a given property. The sliced program 
is then passed to the symbolic execution engine for path-based exploration.

%=======================================================================================
%\subsection{\verifox: Path-wise Symbolic Execution Tool} 
%=======================================================================================
%\Omit{
%\begin{figure}[tp]
%\centering
%\vspace*{0.3cm}
%\scalebox{.50}{\import{figures/}{work-flow.pspdftex}}
%\caption{Automated Slicing and Trace partitioning\label{fig:desc}}
%\end{figure}
%}
\Omit{
\begin{figure}[t]
\tikzstyle{rblock}=[rounded corners,draw=black]
\tikzstyle{block}=[draw=black]
\begin{tikzpicture}

@@

\end{tikzpicture}
\caption{Automated Slicing and Trace partitioning\label{fig:desc}}
\end{figure}
}
%
\input{verifox}
%
\subsection{Approach 2: Monolithic Co-verification Technique}
%
\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.6]{figures/traditional_flow.pdf}
\caption{Conventional HW/SW Co-verification Flow}
\label{fig:conventional}
\end{center}
\end{figure}

Figure~\ref{fig:conventional} illustrates the conventional approach to HW/SW
co-veri\-fi\-cation, as implemented in the state-of-the-art co-verification
tool, \hwcbmcv~\cite{CKY03,vlsid13}.  \hwcbmcv is the only open-source
formal co-verification tool that supports Verilog RTL (System Verilog 2009
standard) and C/C++ firmware.  In contrast to our proposed approach,
\hwcbmcv maintains two separate flows for HW and FW.  The top flow in
Figure~\ref{fig:conventional} uses synthesis to obtain either a bit-level or
a word-level netlist from Verilog RTL.  The bottom flow illustrates the
translation of the C program into static single assignment (SSA) form. 
These two flows meet only at the solver.  Thus, \hwcbmcv generates a
monolithic formula from the C and RTL description, which is then checked
with SAT/SMT solvers.  \hwcbmcv provides specific handshake primitives such
as $next\_timeframe()$ and $set\_inputs()$ to model FW-HW communications.
%
\Omit{
The hardware and firmware run independently of each other. 
The communication between them takes place through these function calls. 
}
%===============================================================================
\section{Experimental Results}
%===============================================================================
